---
title: "Ebola in Guinea, Sierra Leone, Senegal"
author: "Salome Hauselmann"
date: today
format: 
  html:
    toc: true
    toc-location: body
    code.fild: show
    code-tools: true
    df-print: paged
    embeded-resources: true

  pdf: 
    toc: true
  docx: 
    toc: true
---

```{r}
#| output: false
#| echo: false
# Load libraries
library(here)
library(tidyverse)
library(lubridate)
library(unibeCols)
```

## Introduction

The aim of this document is to describe the increase in Ebola cases and deaths in Guinea, Sierra Leone, and Senegal in 2014 and 2015.

## Data

We read and process the data of Ebola cases and deaths in Guinea, Sierra Leone, and Senegal as reported by the [Projects in R Course GitHub repository](https://github.com/ISPMBern/projects-in-R).

```{r}
#| output: false
#| echo: false
# Read data
ebola <- read_csv(here("data/raw/ebola.csv"))
data_ebola_cum_cases <- ebola %>% 
  select(date = Date, country = Country, Cum_conf_cases = Cum_conf_cases, Cum_conf_death = Cum_conf_death) %>% 
  filter(date <= ymd("2015-12-29") & 
        (country == "Guinea" | country ==  "Sierra Leone" | country == "Senegal"))
```

@fig-cases1 shows the number of Confirmed Cases of Ebola in Guinea, Sierra Leone and Senegal from 2014 to 2015.

```{r}
#| label: fig-cases1
#| fig-cap: "Confirmed Cases of Ebola in Guinea, Sierra Leone and Senegal."
#| fig_width: 8
#| fig_height: 4
ggplot(data = data_ebola_cum_cases, 
               mapping = aes(x = date, y = Cum_conf_cases, colour = country)) + 
  geom_line(mapping = aes(group = country), 
            alpha = 0.7, linetype = "dashed", linewidth = 1.5) +
  scale_colour_manual(name = "Country",
                      breaks = c("Guinea", "Sierra Leone", "Senegal"),
                      values = c(unibeRedS()[1], unibeOceanS()[1], unibeMustardS()[1]),
                      labels = c("GIN", "SLE", "SEN" )) +
  ggtitle(label = "Confirmed Cases of Ebola") +
  xlab(label = "Time") +
  ylab(label = "Cum. # of confirmed cases")
```

@fig-cases2 shows the number of confirmed Ebola death in Guinea, Sierra Leone and Senegal 2014 to 2015.

```{r}
#| label: fig-cases2
#| fig-cap: "Ebola deaths confirmed in Guinea, Sierra Leone and Senegal."
#| fig_width: 8
#| fig_height: 4
ggplot(data = data_ebola_cum_cases, 
               mapping = aes(x = date, y = Cum_conf_death, colour = country)) + 
  geom_line(mapping = aes(group = country), 
            alpha = 0.7, linetype = "dashed", linewidth = 1.5) +
  scale_colour_manual(name = "Country",
                      breaks = c("Guinea", "Sierra Leone", "Senegal"),
                      values = c(unibeRedS()[1], unibeOceanS()[1], unibeMustardS()[1]),
                      labels = c("GIN", "SLE", "SEN" )) +
  ggtitle(label = "Confirmed Death of Ebola") +
  xlab(label = "Time") +
  ylab(label = "Cum. # of confirmed death")
```


## Conclusion

The largest increase in the number of Ebola cases as well as the number of Ebola deaths can be seen in Guinea compared to Sierra Leone and Senegal.

